---
title: "607 Final Project"
author: "Weijian Cai, Qixing Li"
date: "May 6, 2019"
output: 
  html_document:
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float:
      collapsed: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE)
```

#Data

##Data Source

We obtained data from [Kaggle.com](https://www.kaggle.com/wsj/college-salaries) 


```{r}
# load packages and data
library(tidyverse)
library(plotly)
library(stringr)
library(magrittr)
library(DT)
```


##Data Load

We read data from the CSV file which was uploaded to Github.
```{r}
# salaries by degree
url_deg <- "https://raw.githubusercontent.com/qixing810/CUNYSPS-DataScience/master/DS607/Project/final%20project/degrees-that-pay-back.csv"
df_deg <- read_csv(url_deg,
                   col_names = c("major", "start_med_slry", "mid_car_slry",
                                 "percent_chng", "mid_car_10th", "mid_car_25th",
                                 "mid_car_75th", "mid_car_90th"),  # rename the column names
                   skip = 1)  # names specified, skip header

# salaries by college type
url_col <- "https://raw.githubusercontent.com/qixing810/CUNYSPS-DataScience/master/DS607/Project/final%20project/salaries-by-college-type.csv"
df_col <- read_csv(url_col,
                   col_names = c("school_name", "school_type", "start_med_slry",
                                 "mid_car_slry", "mid_car_10th", "mid_car_25th",
                                 "mid_car_75th", "mid_car_90th"),
                   skip = 1)

# salaries by region
url_reg <- "https://raw.githubusercontent.com/qixing810/CUNYSPS-DataScience/master/DS607/Project/final%20project/salaries-by-region.csv"
df_reg <- read_csv(url_reg,
                   col_names = c("school_name", "region", "start_med_slry",
                                 "mid_car_slry", "mid_car_10th", "mid_car_25th",
                                 "mid_car_75th", "mid_car_90th"),
                   skip = 1)
```


#Tidy and Wrangle

##Data Structure
First we looked at the structure of the dataset.
```{r}
str(df_deg)
str(df_col)
str(df_reg)
```

##Data Types
We removed the $ in the numbers and changed the data types from character to numeric in each dataset.
```{r}
# salaries by degree
df_deg$start_med_slry <- parse_number(df_deg$start_med_slry)
df_deg$mid_car_slry <- parse_number(df_deg$mid_car_slry)
df_deg$mid_car_10th <- parse_number(df_deg$mid_car_10th)
df_deg$mid_car_25th <- parse_number(df_deg$mid_car_25th)
df_deg$mid_car_75th <- parse_number(df_deg$mid_car_75th)
df_deg$mid_car_90th <- parse_number(df_deg$mid_car_90th)
datatable(df_deg)
```


```{r}
# salaries by college type
df_col$start_med_slry <- parse_number(df_col$start_med_slry)
df_col$mid_car_slry <- parse_number(df_col$mid_car_slry)
df_col$mid_car_10th <- parse_number(df_col$mid_car_10th)
df_col$mid_car_25th <- parse_number(df_col$mid_car_25th)
df_col$mid_car_75th <- parse_number(df_col$mid_car_75th)
df_col$mid_car_90th <- parse_number(df_col$mid_car_90th)
datatable(df_col)
```

```{r}
# salaries by region
df_reg$start_med_slry <- parse_number(df_reg$start_med_slry)
df_reg$mid_car_slry <- parse_number(df_reg$mid_car_slry)
df_reg$mid_car_10th <- parse_number(df_reg$mid_car_10th)
df_reg$mid_car_25th <- parse_number(df_reg$mid_car_25th)
df_reg$mid_car_75th <- parse_number(df_reg$mid_car_75th)
df_reg$mid_car_90th <- parse_number(df_reg$mid_car_90th)
datatable(df_reg)
```

##data transformation
start salary vs mid career salaries distribution analysis
```{r}
start_vs_mid <- df_reg%>%
                 select(start_med_slry, mid_car_slry)%>%
                 gather(career, salary)%>%
                 mutate(career =as_factor(career,fct_rev))
head(start_vs_mid)
```
```{r}
ggplot(start_vs_mid, aes(salary, fill=career))+
  geom_histogram(position ="dodge")
  
```

#need to change to numeric, find median of starting and mid salary
```{r}
start_mean<- start_vs_mid%>%
             filter(career=="start_med_slry")%>%
             median()
```

The distribution for starting median salary is definitely concentrated at the lower range of salaries and is right-skewed. Graduates of most schools start out with median of $45,000. Towards mid-career, the distribution of median salaries becomes more dispersed and the median increases to $80,000.

##scatter plot
correlation between starting and mid-career salaries?
```{r}
ggplot(df_reg, aes(start_med_slry,mid_car_slry))+
  geom_point()+
  geom_smooth()
```

There is a strong positive correlaton between the starting salaries and the mid-career salaries.

##salary by region
##salary by college
##salary by majores
   starting median salary by major
   starting and mid career median salary by major

   